---
title: "Labour Market Resilience"
subtitle: "Economic Crisis Has Diverse Impacts on Industry Work Hours"
author: 
  - Mingjia Chen
  - Catherine Punnoose
  - Tianen (Evan) Hao
thanks: "Code and data are available at: https://github.com/ShadyEvan4830/Labour_Market_Resilience.git"
date: March 11, 2024
date-format: long
abstract: "First sentence, second sentence, third sentence."
format: pdf
toc: true
number-sections: true
bibliography: bibliography.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

#download the packages if necessary, then load the packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(janitor,here,dplyr,tidyverse,knitr,kableExtra,ggplot2)

library(janitor)
library(tidyverse)
library(knitr)
library(dplyr)
library(kableExtra)
library(here)
library(ggplot2)
```

# Introduction

# Data

## Source Data

For the analysis, we retrieved the following data as described in @tbl-use-var.

| Variable | New Name   | Description                     | Example |
|----------|------------|---------------------------------|---------|
| response | work hours | Respondent's response           | 23      |
| year     | year       | The year of the survey recorded | 2007    |

: Source data retrieved from GSS {#tbl-use-var}

## Data Cleaning

The data was downloaded and filtered for the selected variables from the selected data variables from GSS[^1]. The data cleaning was performed based on value definitions as defined in the GSS codebook [@gssCodebook]. One of the variable names `response` is renamed to be more informative [@tbl-use-var].

The data was thus been cleaned by using the open source statistically programming language R [@citeR], with libraries `tidyverse` [@tidyverse], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `readr` [@readr], `tibble` [@tibble], `here` [@here], `kableExtra` [@kableExtra], `janitor` [@janitor], and `knitr` [@knitr].

[^1]: https://gss.norc.org/documents/stata/GSS_stata.zip

```{r}
#| label: tbl-data-overview
#| tbl-cap: Overview of the Categorized Data 
#| echo: false
#| warning: false
#| message: false

data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
# Delete the column
data <- data[c(2:ncol(data))]
colnames(data)[1] <- "Work Hours/Years"
# Get rid of the first letter X for each of the years
colnames(data) <- gsub("X", "", colnames(data))

data_selected <- data[c("Work Hours/Years", "1998","2008","2021","Total")]
kable(data_selected,row.names = FALSE)|>
  kable_styling() |>
  row_spec(6, hline_after = TRUE)
```

## Data Limitations 

## Survey Methodology 

### 1998 Survey Methodology 
```{r}
#| label: fig-Survey-1998
#| tbl-cap: Overview of the 1998 Survey Data 
#| echo: false
#| warning: false
#| message: false
data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[1:6,]

ggplot(data, aes(x = "", y = X1998, fill = work_hours)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Set3") +  # Use a color palette
  theme_void() +  # Remove background and gridlines
  labs(title = "1998 Survey Results") +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```
\break

### 2008 Survey Methodology

```{r}
#| label: fig-Survey-2008
#| tbl-cap: Overview of the 2008 Survey Data 
#| echo: false
#| warning: false
#| message: false

data <-read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[1:6,]

ggplot(data, aes(x = "", y = X2008, fill = work_hours)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Set3") +  # Use a color palette
  theme_void() +  # Remove background and gridlines
  labs(title = "2008 Survey Results") +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```
\break

### Gap in 2020 Data 
```{r}
#| label: fig-Survey-2021
#| tbl-cap: Overview of the 2021 Survey Data 
#| echo: false
#| warning: false
#| message: false
data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[1:6,]

ggplot(data, aes(x = "", y = X2021, fill = work_hours)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_brewer(palette = "Set3") +  # Use a color palette
  theme_void() +  # Remove background and gridlines
  labs(title = "2021 Survey Results") +
  theme(plot.title = element_text(hjust = 0.5, size = 16))
```

### Non-responses


# Results

## Working Hour Comparison Between Financial Events 

```{r}
#| label: fig-ave-hour-comparison
#| fig-cap: Average Working Hours Comparison between Financial Events
#| echo: false
#| warning: false
#| message: false

data <- read.csv(here::here("inputs/data/cleaned_data.csv"))
# Get rid of the first letter X for each of the years
colnames(data)[1] <- "work_hours"
colnames(data) <- gsub("X", "", colnames(data))

#Histogram of average working hours for 2008, 2021, 2022, and Total
hist_data <- select(data, work_hours,"2010","2018","2021")
hist_data <- hist_data[5:94,]

num <- 3 #number of years
sum <- rep(0,num)
total_people <- rep(0,num)
for(i in 1:num){
  total_people[i] <- sum(hist_data[, i+1])
}
hist_data[, 1] <- sapply(hist_data[, 1], as.numeric)
hist_data[90,1] <- 90

sum[1] <- sum(hist_data$"2010" * hist_data$work_hours)
sum[2] <- sum(hist_data$"2018" * hist_data$work_hours)
sum[3] <- sum(hist_data$"2021" * hist_data$work_hours)


averages <- round(sum/total_people) 
years <- c("2010","2018","2021")

average_hours <- data.frame(cbind(averages,years))

ggplot(average_hours,aes(x=years,y=averages)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Year", y = "Average Working Hour") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

\break 

```{r}
#| label: fig-cate-comparisons
#| fig-cap: Categorical Working Hour Comparison between the Years
#| fig-subcap: ["Asian Financial Crisis (1998)","World Financial Crisis(2008)","During Covid-19 Pandemic (2021)", "Total"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

#still need to modify the graphs

data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[c(2:ncol(data))]
# Get rid of the first letter X for each of the years
colnames(data) <- gsub("X", "", colnames(data))

Data_1998 <- select(data,work_hours,`1998`)
ggplot(Data_1998,aes(x=work_hours,y=`1998`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2008 <- select(data,work_hours,`2008`)
ggplot(Data_2008,aes(x=work_hours,y=`2008`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_2021 <- select(data,work_hours,`2021`)
ggplot(Data_2021,aes(x=work_hours,y=`2021`)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

Data_Total <- select(data,work_hours,Total)
ggplot(Data_Total,aes(x=work_hours,y= Total)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Working Hours", y = "Number of Reports") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

\break

## Nonreponse Rate
```{r}
#| label: fig-nonresponse-rate
#| fig-cap: Nonresponse Rates
#| echo: false
#| warning: false
#| message: false

# Calculate percentages
data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[c(2:ncol(data))]
data <- data[1:6,]

#colnames(data) <- paste(colnames(data),"Year",sep="_") 

colnames(data)<- as.character(colnames(data))

#data_long <- data %>% 
#  pivot_longer(cols = everything(), names_to = "WorkHours", values_to #= "Response")

# Select the 'No Response' row and remove the first column which is not needed for the plot
no_response_data <- data[data$work_hours == 'No Response', -1]

# We also need to remove the 'Total' column as we did in the Python example
no_response_data <- no_response_data[-length(no_response_data)]

# Calculate total responses per year excluding the 'work_hours' and 'Total' columns

data <- read.csv(here::here("inputs/data/cleaned_categorized_data.csv"))
data <- data[c(3:ncol(data))]

total_responses_per_year <- as.vector(data[7,])
total_responses_per_year <- unlist(total_responses_per_year)

# Calculate non-response rate as a percentage
non_response_rate <- (no_response_data / total_responses_per_year) * 100

years <- colnames(non_response_rate)
years <- gsub("X", "", years)

non_response_rate <-as.vector(non_response_rate)
non_response_rate <-unlist(non_response_rate)
non_response_data <- data.frame(years,non_response_rate)

# Plot using ggplot2
ggplot(non_response_data, aes(x = years,y=non_response_rate)) +
  geom_bar(stat = "identity",fill = "skyblue") +
  theme_minimal() +
  labs(title = "Non-Response Rate by Year", x = "Year", y = "Percentage of Non-Responses") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```

\break

# Discussion

## Labour Market Response Across Economic Crises

## Demographic Disparities in Labour Market Participation

### Younger Workers and Educational Attainment

### Women in the Labour Market

## Political Affiliation and Labour Market Perceptions

##  Ethical Considerations and Bias

## Limitations

## Future Research

## Journal Study: Directions on Non-response Rates

# Conclusion

\appendix

# Appendix {-}

## Survey Introduction

The supplemental survey is available here: [Google Form Link](https://forms.gle/Ffma5ggEcP6wfumu6)

Welcome to our survey, which builds on our research analyzing the impact of economic downturns on labour dynamics in the U.S. from 1972 to 2022, based on the General Social Survey (GSS). The GSS collects information and historical records of respondents' attitudes and experiences. Our study has highlighted significant shifts in working hours, particularly during crises such as the 2008 Financial Crisis and the COVID-19 pandemic, affecting various demographic groups differently. This survey seeks to deepen our understanding of these impacts and gather more nuanced data on labour force participation across different economic conditions.

Your participation is voluntary and confidential. You may skip any question or withdraw at any time. Responses are anonymized and will be used solely for research purposes. For any inquiries or additional information, please contact us at evan.hao@mail.utoronto.ca, mingjia.chen@mail.utoronto.ca, and/or catherine.punnoose@mail.utoronto.ca.

## Survey Questions

\begin{itemize}
        \item Demographic Information:
            \begin{enumerate}
                \item What is your age group?
                \begin{itemize}
                  \item Under 18
                  \item 18-24
                  \item 25-34
                  \item 35-44
                  \item 45-54
                  \item 55-64
                  \item 65+
                \end{itemize}
                \item What is your highest level of education?
                \begin{itemize}
                  \item High school graduate
                  \item Bachelor's degree
                  \item Graduate or professional degree
                \end{itemize}
                \item Please specify your gender:
                \begin{itemize}
                  \item Male
                  \item Female
                  \item Non-binary
                  \item Prefer not to say
                \end{itemize}
            \end{enumerate}
        \item Employment Status and Industry:
              \begin{enumerate}
                \item What is your current employment status?
                \begin{itemize}
                  \item Employed full-time
                  \item Employed part-time
                  \item Unemployed
                  \item Retired
                  \item Student
                \end{itemize}
                \item If employed, what industry do you work in?
                \begin{itemize}
                  \item Manufacturing
                  \item Construction
                  \item Retail
                  \item Hospitality
                  \item Healthcare
                  \item Education
                  \item Information Technology
                  \item Professional Services
                  \item Other (please specify)
                \end{itemize}
            \end{enumerate}
        \item Impact of Economic Downturns:
              \begin{enumerate}
                \item How did the 2008 financial crisis and/or the COVID-19 pandemic affect your working hours? Did your hours:
                \begin{itemize}
                  \item Increase
                  \item Decrease
                  \item Stay the same
                  \item Not applicable/I wasn't employed during these times
                \end{itemize}
                \item Did you have to switch to remote work during the COVID-19 pandemic?
                \begin{itemize}
                  \item Yes
                  \item No
                  \item Not applicable
                \end{itemize}
              \end{enumerate}
        \item Responses to Economic Downturns:
              \begin{enumerate}
                \item In times of economic downturn, what support measures do you believe are most effective for workers?
                \begin{itemize}
                  \item Unemployment benefits
                  \item Skills retraining programs
                  \item Direct financial aid
                  \item Flexible work arrangements
                  \item Other (please specify)
                \end{itemize}
            \end{enumerate}
\end{itemize}

## Survey Completion Message

Thank you for participating in our survey. Your insights are invaluable to our ongoing research into labour market dynamics and the effects of economic downturns on working hours. Your contribution helps us identify areas for targeted interventions and policy recommendations to support vulnerable labour market groups.
\newpage
# References